Stability of variational quantum dynamics with Gaussian approximants

### 2013-07-06 paper.tex
writing

\input a4
\input xpmath
\input respnotes

@@

This claims that $G$ is a projection: if the inner product is normalised properly, $G$ keeps the component of $f$ in the space of entire functions, and discards the orthogonal component.  In that case, we'd have a simple underdetermined problem, which could be regularised very easily by discarding the small singular values.  But that isn't what I see with discrete versions: they seem to have a steady decay of singular values, not a sudden drop to zero.  Is that an artefact of how I'm discretising?

@@

>>>>

\nonstopmode
\input a4
\input xpmath
\input respnotes

@@

So $G$ is a projection: if the inner product is normalised properly, $G$ keeps the component of $f$ in the space of entire functions, and discards the orthogonal component.  The continuous problem is simply underdetermined, and the inverse problem can be regularised very easily by excluding components with zero singular values from the solution.  That isn't what I see with discrete versions: they seem to have a steady decay of singular values, not a sudden drop to zero.  This must be entirely an artefact of the discretisation.

This paper will talk about two related but distinct inverse problems.  Both start with a given wavefunction $|ψ\rangle$, and a set of amplitudes in phase space, $\{α_j\}$, over which to expand it.  In systems with few modes, these points could be a grid, but in many-mode systems they will have to be more sparse than that: there isn't space to store grid coefficients.  The first problem, let's call it the inverse problem, is to find a set of weights $f_j$ such that $∑f_j|α_j\rangle$ approximates $|ψ\rangle$.  This is simply finding a discrete inverse to $G$ sampled on the grid $α_j$; since $G$ is a projection, the pseudoinverse of $G$ should solve it well enough.  The differential case, where we have an expansion for $|ψ\rangle$ which we want to perturb to represent $|ψ\rangle+|dψ\rangle$, is no harder.  Because $G$ is linear, we can expand $|dψ\rangle$, and add it to the expansion of $|ψ\rangle$.

The second problem, the variational one, is a generalisation.  As well as the grid $α_j$, we have a set of weights $f_j$ representing some wavefunction, presumably not $|ψ\rangle$.  Instead of expanding $|ψ\rangle$ over phase space, we want to expand it over the columns of $|Dψ(f,α)\rangle$, the set of all changes we can make by slightly adjusting the original expansion, and the grid.  Obviously, we can solve this problem simply by setting the coefficients of $|∂fψ(f,a)\rangle$ to the solution of the inverse problem, and those of $|∂fψ(f,a)\rangle$ to zero.  Is this the smallest norm solution?  Probably not: if a coherent state was moving in an oscillator, it might be easier to just rotate the $α_j$.

Since we have a totally stable solution to the problem, why would we try anything else?  Because we're iterating the problem, and we need to preserve our ability to solve it at future steps.  If the amplitudes form a grid, which covers all the energies the system could possibly access, we're set: we'll always be able to expand over them.  However, we can't do that for many mode systems: there will always be accessible states of the system whose support in phase space is remote from the current grid.  The set $|Dψ\rangle$ will have some small singular values, corresponding to directions where the amplitudes and samples are adjusted in a way that preserves the wavefunction.  We are free to add these as we like, in order to maintain a grid on which we can expand $|ψ\rangle$ in the future.

So what are the critera for adjusting the grid?  I think there are two obvious ones: firstly, that the inverse problem stays well-posed.  Otherwise, we're wasting grid points.  Secondly, when we're expanding a derivative over the tangent space $|Dψ\rangle$, we want to tilt the tangent space closer to the derivative.  That's a second order effect, and won't come naturally from a least squares adjustment of $f$ and $α$.  The first criterion seems redundant: if we could better represent the goal by shifting two sample points apart, the second criterion will do that automatically.  However, if we only care about the tangent space and not the derivative map, we might need very large adjustments of the amplitudes and weights to get to the point in the tangent space that we want.  So a condition number criterion seems sensible.  There might be other criteria that I've overlooked.

I still haven't thought in detail about the results with truncated Fock spaces.  Obviously, when we restrict the domain of $G$ to the lowest $N$ energy levels, the restricted operator has at most $N$ singular directions.  Since the low energy levels are the ones a restricted grid tends to represent, those singular directions will have relatively large singular values.  So when we do this, we get a well-posed problem, but we lose the ability to add components with small singular values, in order to expand the states we'll get in the future.  And in fact, by solving the inverse problem, we can always get a nice solution in any case.  Numerical experiments suggest that the variational problem is quite well posed for a single step, too.

@@

It turns out that complex convolutions are not like real convolutions: the imaginary part of $t-t'$ can control the rate at which the kernel oscillates.  Each line of constant imaginary part in the convolution is determined by a limited range of frequencies from the function, but that frequency range can vary with the imaginary part, so that all frequency components of the function contribute to the convolution somewhere.  Is the Fourier transform a complex convolution?

Not a
